---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)

library(magrittr)
devtools::load_all()
```

# pformat
pformat provides powerful string interpolation and formatting capabilities. It
is a R language implementation of Python's new style string formatting API, 
coupled with some original features.

## Features

* Supports numeric (`integer`, `double`, `complex`), date (`Date`, 
`IDate`, `POSIXt`) and `character` types out-of-the-box.
* Supports **expressions** inside fields, avoiding intermediate variables.
* named fields can be evaluated inside lists or data frames, saving typing.
* Supports evaluation on the current environment. You can feed `pformat()` only
the format string, and it will look for corresponding data on the environment.
* It is **vectorized**, allowing formatting of extensive amounts of data with 
a single call.
* It is possible to **preparse** format strings, which may avoid unnecessary 
calls and reduce computing time inside loops.
* It is **extensible**: you can write custom formatters for your classes.

## Examples

### Basic usage

A pair of braces and everything inside them form a placeholder. The simplest use 
case is that of positional formatting: arguments are assigned to placeholders 
according to their position.

```{r}
pformat("{} {}", "one", "two")
```

You can give placeholders an explicit positional index. This allows for
re-arranging the order of display without changing the arguments. *Note:* 
contrary to Python, in pformat indices start from 1 and not from 0.

```{r}
pformat("{2} {1}", "one", "two")
```

### Vectorization

pformat is vectorized, so you can generate many formatted strings using
a single call

```{r}
pformat("Name: {}; Age: {}", c("Abby", "Bob", "Carl"), 22:24)
```

Usual recycling rules apply:

```{r}
pformat("{}-{}{}", "expr", c("a", "b", "c"), 1:2)
```

### Named placeholders

pformat provides three ways of using named placeholders: 

* keyword arguments.
* passing a list, data frame, or environment as the first parameter.
* evaluation on the current environment.

That's also the order which pformat uses when looking for corresponding names. 
The following example illustrates the three methods, where all calls produce 
same output.

```{r}
# keyword arguments
pformat("Name: {name}; Age: {age}", name = c("Abby", "Bob", "Carl"), age = 22:24)

# passing a data frame as the first parameter
people = data.frame(name = c("Abby", "Bob", "Carl"), age = 22:24)
pformat(people, "Name: {name}; Age: {age}")

# evaluation on the environment
name = c("Abby", "Bob", "Carl")
age = 22:24
pformat("Name: {name}; Age: {age}")
```

### Piping

Passing data frames as the first parameter allows pformat to be chained using the pipe operator `%>%` provided by package `magrittr`.

### Expressions

Placeholders can hold not only identifiers but any R expression, provided
that its result type is supported by pformat.

```{r}
n = 9
pformat("{n} x {i} = {n * i}", i = 1:10)
```

```{r}
df = data.frame(name = c("Walter", "Frederick", "Lindsey"), 
           surname = c("Unzueta", "Winstead", "Chambers"))
pformat(df, "{substr(name, 1, 1)}. {surname}")
```

## Formatting

### Integers, floating point numbers, and strings

pformat uses (will use) the same format string syntax as Python. See 
[Python docs](https://docs.python.org/3/library/string.html#string-formatting). 
Development in progress.

### Dates

Date formatting uses the traditional `strftime()` conversion specification, just 
like Python. See `strftime`'s help page for more details.

```{r}
pformat("Mother's day: {:%d/%m/%Y}", as.Date("2016-05-08"))
```